% --------------------------------------------------
% Abstract (Bilingual: English / German) -- REVISED
% --------------------------------------------------

\begin{abstract}

\textbf{English Version}

We investigate Topological Integrity Gravity (TIG) as an effective extension of General Relativity motivated by spectral considerations and constrained by an additional stability functional.

In the low-curvature regime, TIG reduces to a quadratic $f(R)$-type theory, reproducing the structure of Starobinsky gravity. In contrast to phenomenological $f(R)$ models, the form of curvature corrections is not arbitrary but restricted by the underlying construction, leading to a reduced and structurally controlled parameter space.

The model introduces an additional scalar curvature degree of freedom associated with the quadratic correction, with a characteristic mass scale determined by the fundamental parameters of the theory.

We analyze a class of static, spherically symmetric configurations exhibiting regular de Sitter-like cores and non-trivial horizon structures. We identify a structurally controlled critical transition in horizon formation governed by a dimensionless parameter $\beta$, which separates horizon-bearing from horizonless compact configurations. This transition is associated with a degeneration of the horizon equation and induces a non-analytic change in the global causal structure.

Near the critical regime, the metric function develops an extended near-zero region, leading to strong modifications of null geodesic propagation and a divergence of echo delay times. This behavior provides a geometric mechanism for long-lived gravitational-wave echoes without requiring exotic boundary conditions.

While the resulting geometries are kinematically regular and thermodynamically consistent, a complete dynamical stability analysis remains an open problem.

TIG therefore provides a constrained and structurally grounded framework for curvature corrections beyond phenomenological $f(R)$ gravity, linking geometric structure, horizon formation, and stability within a unified effective description.

\vspace{1em}

\textbf{Deutsche Version}

Wir untersuchen Topological Integrity Gravity (TIG) als eine effektive Erweiterung der Allgemeinen Relativitätstheorie, die durch spektrale Überlegungen motiviert und durch ein zusätzliches Stabilitätsfunktional eingeschränkt ist.

Im Niedrigkrümmungsregime reduziert sich TIG auf eine quadratische $f(R)$-artige Theorie und reproduziert die Struktur der Starobinsky-Gravitation. Im Gegensatz zu phänomenologischen $f(R)$-Modellen ist die Form der Krümmungskorrekturen nicht frei wählbar, sondern durch die zugrunde liegende Konstruktion eingeschränkt, was zu einem reduzierten und strukturell kontrollierten Parameterraum führt.

Das Modell führt einen zusätzlichen skalaren Krümmungsfreiheitsgrad ein, der mit der quadratischen Korrektur verknüpft ist und eine charakteristische Massenskala besitzt, die durch die fundamentalen Parameter der Theorie bestimmt wird.

Wir analysieren eine Klasse statischer, sphärisch symmetrischer Konfigurationen mit regulären de-Sitter-artigen Kernen und nicht-trivialen Horizontstrukturen. Dabei identifizieren wir einen strukturell kontrollierten kritischen Übergang in der Horizontbildung, der durch einen dimensionslosen Parameter $\beta$ gesteuert wird und Konfigurationen mit und ohne Horizont trennt. Dieser Übergang ist mit einer Degeneration der Horizontgleichung verbunden und führt zu einer nicht-analytischen Änderung der globalen kausalen Struktur.

In der Nähe des kritischen Regimes entwickelt die Metrikfunktion eine ausgedehnte Near-Zero-Region, was zu starken Modifikationen der Nullgeodäten und zu einer Divergenz der Echo-Verzögerungszeiten führt. Dieses Verhalten liefert einen geometrischen Mechanismus für langlebige Gravitationswellen-Echos ohne exotische Randbedingungen.

Während die resultierenden Geometrien kinematisch regulär und thermodynamisch konsistent sind, bleibt eine vollständige dynamische Stabilitätsanalyse eine offene Aufgabe.

TIG stellt somit einen eingeschränkten und strukturell fundierten Rahmen für Krümmungskorrekturen jenseits phänomenologischer $f(R)$-Modelle dar und verbindet geometrische Struktur, Horizontbildung und Stabilität in einer einheitlichen effektiven Beschreibung.

\end{abstract}
