\documentclass[11pt,a4paper]{article}

% ------------------------------------------------------------
% Encoding and language
% ------------------------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% ------------------------------------------------------------
% Mathematics
% ------------------------------------------------------------
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

% ------------------------------------------------------------
% Layout and links
% ------------------------------------------------------------
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{enumitem}

\geometry{margin=2.5cm}

% ------------------------------------------------------------
% Theorem environments
% ------------------------------------------------------------
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}

% ------------------------------------------------------------
% Metadata
% ------------------------------------------------------------
\title{Topological Integrity Gravity: A Structural Horizon Transition}
\author{Kai Stefan Dietrich}
\date{Version 1.0}

\begin{document}

\maketitle

\begin{abstract}
We introduce Topological Integrity Gravity (TIG), a structurally constrained
framework for gravitational horizon formation. The central idea is that admissible
curvature corrections should not be treated as arbitrary phenomenological
deformations of General Relativity, but as structures constrained by consistency
requirements. Within a minimal static and spherically symmetric reduction, these
conditions lead to a cubic normal form governing horizon formation. The resulting
horizon equation exhibits a critical transition separating configurations with
horizons from compact configurations without horizons. This transition is
controlled by a dimensionless parameter and is characterized by the coalescence
of roots in the associated cubic equation. We derive the critical parameter,
analyze the root structure, and discuss the interpretation of the transition as
a structural threshold rather than a continuous perturbative deformation.
\end{abstract}

\tableofcontents

\section{Introduction}

General Relativity provides the most successful classical description of
gravitation. Its geometric formulation identifies gravitational interaction
with the curvature of spacetime and has passed a wide range of observational
tests. Nevertheless, the appearance of singularities and the structure of
horizon formation indicate that the classical theory may not fully encode all
structural constraints relevant at extreme curvature scales.

A common strategy is to modify the gravitational action by adding curvature
corrections. Such models are often written as extensions of the Einstein-Hilbert
action through a generalized function \(f(R)\) of the Ricci scalar. However, this
procedure may introduce an overly large space of possible theories. If curvature
corrections are chosen freely, the resulting model space becomes
phenomenological rather than structurally constrained.

The purpose of this work is to formulate a minimal alternative viewpoint. We
assume that admissible curvature corrections are restricted by consistency
conditions. In this sense, the question is not merely which correction terms can
be written down, but which correction terms preserve the structural integrity of
horizon formation.

We call this framework \emph{Topological Integrity Gravity} (TIG). The term
``topological integrity'' is used here in a restricted mathematical sense: it
refers to the preservation of the qualitative root structure associated with
horizon formation under admissible curvature corrections.

The central result of this paper is the emergence of a cubic normal form,
\begin{equation}
x^3 - x^2 + \beta^3 = 0,
\label{eq:intro_cubic}
\end{equation}
where \(x\) is a dimensionless radial horizon variable and \(\beta\) is a
dimensionless structural control parameter. This equation exhibits a critical
transition: for sufficiently small \(\beta\), horizon solutions exist; beyond a
critical value, no positive horizon root remains.

The transition is not introduced by hand. It follows from the algebraic structure
of the reduced horizon equation. The resulting critical point is therefore
interpreted as a structural boundary between horizon and non-horizon
configurations.

\section{Geometric and Physical Setup}

We consider a four-dimensional spacetime \((M,g)\) equipped with a Lorentzian
metric \(g_{\mu\nu}\). The discussion is restricted to static, spherically
symmetric configurations. The general line element may be written in the form
\begin{equation}
ds^2 = -A(r)c^2dt^2 + B(r)dr^2 + r^2 d\Omega^2,
\label{eq:metric}
\end{equation}
where \(d\Omega^2\) denotes the metric on the unit sphere.

In the Schwarzschild solution of General Relativity, the horizon is located at
the radial coordinate
\begin{equation}
r_s = \frac{2GM}{c^2},
\label{eq:schwarzschild_radius}
\end{equation}
where \(M\) is the mass of the central object.

The horizon condition can be expressed through the vanishing of the lapse
function,
\begin{equation}
A(r)=0.
\label{eq:horizon_condition}
\end{equation}

In modified or structurally constrained theories, the functional form of
\(A(r)\) may receive corrections. The key question is whether such corrections
preserve a well-defined horizon structure.

\begin{definition}[Dimensionless horizon variable]
We define the dimensionless radial variable
\begin{equation}
x := \frac{r}{r_s}.
\label{eq:x_definition}
\end{equation}
The Schwarzschild horizon corresponds to \(x=1\).
\end{definition}

\begin{definition}[Structural control parameter]
Let \(\ell\) denote a characteristic correction length scale associated with the
regularization or structural modification of the central region. We define the
dimensionless parameter
\begin{equation}
\beta := \frac{\ell}{r_s}.
\label{eq:beta_definition}
\end{equation}
The regime \(\beta \ll 1\) corresponds to corrections much smaller than the
Schwarzschild scale.
\end{definition}

The limit \(\beta=0\) must recover the classical Schwarzschild horizon. This
requirement imposes a normalization condition on any admissible horizon equation:
the dimensionless equation must admit \(x=1\) as the physical root when
\(\beta=0\).

\section{Structural Consistency Principle}

The guiding assumption of TIG is that gravitational corrections are not arbitrary
deformations. Instead, they must preserve the qualitative integrity of the
horizon formation problem.

\begin{definition}[Structural horizon consistency]
A horizon equation is structurally consistent if:
\begin{enumerate}[label=(\roman*)]
\item it admits the Schwarzschild limit \(x=1\) when the correction parameter
vanishes;
\item it has a well-defined algebraic root structure;
\item horizon disappearance, if it occurs, takes place through a critical root
coalescence rather than by arbitrary removal of solutions;
\item the correction terms are dimensionless and compatible with the chosen
normalization.
\end{enumerate}
\end{definition}

These conditions are minimal. They do not specify a complete gravitational field
theory. Instead, they constrain the admissible reduced horizon equations.

The simplest nontrivial algebraic setting in which a horizon can disappear
through root coalescence is a cubic equation. A quadratic equation can describe
a pairwise disappearance of roots, but does not naturally encode the
Schwarzschild root together with a degenerate correction sector. The cubic
normal form is therefore the minimal algebraic structure capable of encoding:
\begin{enumerate}[label=(\alph*)]
\item the Schwarzschild root at \(x=1\),
\item a degenerate zero-radius sector,
\item a finite correction scale.
\end{enumerate}

\section{Derivation of the Cubic Normal Form}

We begin with the degenerate Schwarzschild-compatible polynomial
\begin{equation}
P_0(x) = x^2(x-1).
\label{eq:degenerate_polynomial}
\end{equation}
This polynomial has a simple physical root at \(x=1\) and a double root at
\(x=0\). The root at \(x=1\) represents the classical horizon, while the double
root at the origin represents a degenerate internal sector.

Expanding \eqref{eq:degenerate_polynomial}, we obtain
\begin{equation}
P_0(x)=x^3-x^2.
\label{eq:expanded_p0}
\end{equation}

A finite structural correction must lift the degenerate zero-radius sector while
preserving dimensional consistency. Since \(x\) is dimensionless and \(\beta\)
is dimensionless, the lowest correction compatible with the cubic structure is
proportional to \(\beta^3\). We therefore define
\begin{equation}
P(x,\beta)=x^3-x^2+\beta^3.
\label{eq:normal_form_polynomial}
\end{equation}

The horizon equation is
\begin{equation}
P(x,\beta)=0,
\end{equation}
or explicitly,
\begin{equation}
x^3-x^2+\beta^3=0.
\label{eq:tig_horizon_equation}
\end{equation}

\begin{remark}
Equation \eqref{eq:tig_horizon_equation} is not proposed as a fundamental field
equation. It is a reduced algebraic normal form for the horizon condition under
the structural consistency assumptions stated above.
\end{remark}

\section{Root Structure and Criticality}

The cubic equation
\begin{equation}
x^3-x^2+\beta^3=0
\end{equation}
has a critical point when two real roots coalesce. This occurs when the
polynomial and its derivative vanish simultaneously.

The derivative is
\begin{equation}
\frac{dP}{dx}=3x^2-2x=x(3x-2).
\label{eq:derivative}
\end{equation}

The nonzero stationary point is therefore
\begin{equation}
x_c=\frac{2}{3}.
\label{eq:critical_x}
\end{equation}

At criticality, we impose
\begin{equation}
P(x_c,\beta_c)=0.
\end{equation}
Substituting \(x_c=2/3\), we obtain
\begin{equation}
\left(\frac{2}{3}\right)^3
-
\left(\frac{2}{3}\right)^2
+
\beta_c^3
=0.
\end{equation}

This gives
\begin{equation}
\frac{8}{27}-\frac{4}{9}+\beta_c^3=0,
\end{equation}
and hence
\begin{equation}
\frac{8}{27}-\frac{12}{27}+\beta_c^3=0.
\end{equation}

Therefore,
\begin{equation}
-\frac{4}{27}+\beta_c^3=0,
\end{equation}
so that
\begin{equation}
\beta_c^3=\frac{4}{27}.
\label{eq:beta_c_cubed}
\end{equation}

Thus,
\begin{equation}
\beta_c=\left(\frac{4}{27}\right)^{1/3}.
\label{eq:beta_c}
\end{equation}

\begin{proposition}[Critical horizon transition]
The TIG horizon equation
\begin{equation}
x^3-x^2+\beta^3=0
\end{equation}
admits a critical transition at
\begin{equation}
\beta_c=\left(\frac{4}{27}\right)^{1/3}.
\end{equation}
At this point, two real roots coalesce at
\begin{equation}
x_c=\frac{2}{3}.
\end{equation}
\end{proposition}

\begin{proof}
A double root occurs when \(P(x,\beta)=0\) and \(P'(x,\beta)=0\). Since
\(P'(x)=x(3x-2)\), the nonzero candidate is \(x=2/3\). Substitution into
\(P(x,\beta)=0\) yields \(\beta_c^3=4/27\), completing the proof.
\end{proof}

\section{Discriminant Analysis}

The cubic equation
\begin{equation}
x^3-x^2+\beta^3=0
\end{equation}
can be written in the standard form
\begin{equation}
x^3+ax^2+bx+c=0,
\end{equation}
with
\begin{equation}
a=-1,\qquad b=0,\qquad c=\beta^3.
\end{equation}

The discriminant of a cubic is
\begin{equation}
\Delta
=
a^2b^2
-4b^3
-4a^3c
-27c^2
+18abc.
\label{eq:cubic_discriminant}
\end{equation}

Substituting \(a=-1\), \(b=0\), and \(c=\beta^3\), we find
\begin{equation}
\Delta
=
-4(-1)^3\beta^3
-27\beta^6.
\end{equation}

Thus,
\begin{equation}
\Delta
=
4\beta^3-27\beta^6.
\label{eq:discriminant_beta}
\end{equation}

Factoring,
\begin{equation}
\Delta
=
\beta^3(4-27\beta^3).
\label{eq:discriminant_factored}
\end{equation}

The discriminant vanishes when
\begin{equation}
\beta=0
\end{equation}
or
\begin{equation}
4-27\beta^3=0.
\end{equation}

The nontrivial critical value is therefore
\begin{equation}
\beta_c^3=\frac{4}{27},
\end{equation}
which agrees with \eqref{eq:beta_c_cubed}.

The sign of the discriminant determines the real root structure:
\begin{itemize}
\item \(\Delta>0\): three distinct real roots;
\item \(\Delta=0\): multiple roots;
\item \(\Delta<0\): one real root and two complex conjugate roots.
\end{itemize}

For positive \(\beta\), the critical transition occurs at
\begin{equation}
\beta=\beta_c.
\end{equation}

\section{Physical Interpretation}

The positive roots of the horizon equation correspond to candidate horizon
locations. In the Schwarzschild limit \(\beta=0\), the equation becomes
\begin{equation}
x^3-x^2=0,
\end{equation}
or
\begin{equation}
x^2(x-1)=0.
\end{equation}
The physical nonzero root is
\begin{equation}
x=1.
\end{equation}

For small positive \(\beta\), the physical horizon root is deformed away from
the Schwarzschild value. As \(\beta\) increases, the relevant positive roots
approach each other. At \(\beta=\beta_c\), the roots coalesce at \(x=2/3\).
For \(\beta>\beta_c\), the corresponding pair of real roots no longer exists.

This is the structural horizon transition of TIG. It separates a regime in
which horizon solutions exist from a regime in which compact configurations do
not possess horizons.

The transition is not interpreted as a smooth perturbation of the Schwarzschild
radius. Instead, it represents a qualitative change in the root structure of the
horizon equation.

\section{Comparison with General Relativity}

In General Relativity, the Schwarzschild horizon is located at
\begin{equation}
r=r_s.
\end{equation}
In dimensionless form this is simply
\begin{equation}
x=1.
\end{equation}

In TIG, the horizon condition becomes a cubic equation. The Schwarzschild
solution is recovered in the limit
\begin{equation}
\beta\to 0.
\end{equation}

Thus TIG does not remove the Schwarzschild solution from the theory. Instead,
it embeds it into a larger structural setting in which the existence of horizons
depends on the value of the control parameter \(\beta\).

The classical horizon appears as the limiting member of a family of structurally
constrained configurations.

\section{Scaling Near Criticality}

Near the critical point, root coalescence implies non-linear sensitivity. Let
\begin{equation}
\epsilon = \beta_c-\beta.
\end{equation}
For small positive \(\epsilon\), the separation between the coalescing roots is
expected to scale as a fractional power of \(\epsilon\).

To see this, expand \(P(x,\beta)\) near the critical point:
\begin{equation}
x=x_c+\delta,
\qquad
\beta=\beta_c-\epsilon.
\end{equation}

Since
\begin{equation}
P(x_c,\beta_c)=0
\end{equation}
and
\begin{equation}
P_x(x_c,\beta_c)=0,
\end{equation}
the leading expansion is
\begin{equation}
P(x,\beta)
\approx
\frac{1}{2}P_{xx}(x_c,\beta_c)\delta^2
+
P_{\beta}(x_c,\beta_c)(-\epsilon).
\end{equation}

The derivatives are
\begin{equation}
P_{xx}=6x-2,
\end{equation}
so at \(x_c=2/3\),
\begin{equation}
P_{xx}(x_c)=2.
\end{equation}

Also,
\begin{equation}
P_{\beta}=3\beta^2.
\end{equation}

Thus the leading balance is
\begin{equation}
\delta^2
\sim
3\beta_c^2\epsilon.
\end{equation}

Therefore,
\begin{equation}
\delta
\sim
\epsilon^{1/2}.
\label{eq:sqrt_scaling}
\end{equation}

This square-root scaling is characteristic of saddle-node type criticality.
Observable quantities depending on the horizon root may therefore exhibit
nonlinear scaling near the transition.

\section{Observable Implications}

The TIG transition may affect observables sensitive to horizon-scale geometry.
Possible implications include:

\begin{itemize}
\item modifications of black hole shadow radius;
\item altered photon sphere structure;
\item changes in near-horizon lensing behavior;
\item modified ringdown or echo-like structures near criticality.
\end{itemize}

At the present stage, these implications should be regarded as structural
predictions rather than fully developed observational forecasts. A complete
analysis requires solving the full metric functions and perturbation equations.

Nevertheless, the existence of a critical transition provides a clear target:
observables should exhibit enhanced sensitivity near \(\beta=\beta_c\).

\section{Scope and Limitations}

This paper develops the algebraic horizon structure of TIG. Several important
questions remain open.

First, the complete dynamical field equations associated with the TIG
consistency principle must be formulated. Second, stability of the resulting
compact configurations must be analyzed. Third, observational predictions must
be computed within a fully specified metric model.

The present work should therefore be read as the foundational normal-form
analysis of the TIG horizon transition, not as a complete gravitational theory.

\section{Conclusion}

We have introduced Topological Integrity Gravity as a structurally constrained
approach to horizon formation. Starting from minimal consistency requirements,
we derived a cubic normal form for the horizon equation:
\begin{equation}
x^3-x^2+\beta^3=0.
\end{equation}

This equation exhibits a critical transition at
\begin{equation}
\beta_c=\left(\frac{4}{27}\right)^{1/3},
\end{equation}
where two real roots coalesce at
\begin{equation}
x_c=\frac{2}{3}.
\end{equation}

The resulting structure separates horizon and non-horizon configurations and
suggests that gravitational systems may be constrained by algebraic consistency
conditions beyond ordinary phenomenological deformation.

Future work will focus on deriving the full field equations, analyzing dynamical
stability, and extracting observational signatures.

\section*{References}

\begin{thebibliography}{99}

\bibitem{einstein1915}
A. Einstein,
\textit{Die Feldgleichungen der Gravitation},
Sitzungsberichte der Preussischen Akademie der Wissenschaften, 844--847 (1915).

\bibitem{schwarzschild1916}
K. Schwarzschild,
\textit{On the gravitational field of a mass point according to Einstein's theory},
Sitzungsberichte der K\"oniglich Preussischen Akademie der Wissenschaften (1916).

\bibitem{wald1984}
R. M. Wald,
\textit{General Relativity},
University of Chicago Press (1984).

\bibitem{carroll2004}
S. M. Carroll,
\textit{Spacetime and Geometry: An Introduction to General Relativity},
Addison-Wesley (2004).

\bibitem{faraoni2010}
V. Faraoni and S. Capozziello,
\textit{Beyond Einstein Gravity: A Survey of Gravitational Theories for Cosmology and Astrophysics},
Springer (2010).

\bibitem{sotiriou2010}
T. P. Sotiriou and V. Faraoni,
\textit{f(R) theories of gravity},
Reviews of Modern Physics 82, 451--497 (2010).

\bibitem{nojiri2011}
S. Nojiri and S. D. Odintsov,
\textit{Unified cosmic history in modified gravity: from F(R) theory to Lorentz non-invariant models},
Physics Reports 505, 59--144 (2011).

\bibitem{thorne1994}
K. S. Thorne,
\textit{Black Holes and Time Warps},
W. W. Norton (1994).

\bibitem{chandrasekhar1983}
S. Chandrasekhar,
\textit{The Mathematical Theory of Black Holes},
Oxford University Press (1983).

\bibitem{arnold1992}
V. I. Arnold,
\textit{Catastrophe Theory},
Springer (1992).

\end{thebibliography}

\end{document}
